Modeling the impacts of replacing car trips with combined public transport and cycling: Reproducible methods, results and actionable evidence from biclaR

### 2023-11-30 paper/PaperCEUS/PaperCEUS.Rmd
last revision

---
title: "Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from biclaR"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to public transport were also estimated and monetized.
  The results show that 10% of trips could be made by cycling in combination with public transport. Shifting to cycling for the shorter first and last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the public transport leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

@@

We used 'jittering' to disaggregate the OD data, resulting in a wide spatial distribution trip origins and destinations [@Lovelace2022Jittering].
The method works by sampling 'sub-points' (nodes on the transport network represented in OpenStreetMap in this case) and using these instead of a single point (typically the centroid) to represent trip origins and destinations for each zone.
This method then distributes the trips to desire lines connecting the subpoints based on a 'disaggregation threshold' which determines the maximum number of trips that can be represented by a single desire line.

@@

The jittering pre-processing stage generates a more realistic representation of the trips undertaken than the traditional centroid-based approach but does precisely capture the exactly spatial distribution of trips.
Even where such datasets exist, they cannot be shared for research due to data privacy regulations.

@@

@@

For both *direct* and *safe* route profiles, 10% of the daily trips have the potential to be made by a combination of PT and cycling (up to 5 km on bike). This unveils the potential of cycling as a complementary mode of PT, with the potential to uptake the number of PT trips within the LMA area by as much as 6.3% (in addition to the 825 thousand PT trips reported in the mobility survey).
<!-- place in the end - here is not clear -->

@@

The quantification of benefits can support policy-makers in prioritizing interventions to reduce the reliance on individual motorized transportation.
The presentation of the results in an open access web application will help  communicate decisions.
Furthermore, the provision of datasets resulting from this project provides a foundation for further research and development of new tools and methods.
<!-- Punchy final sentence: -->
The methods are reproducible and based on open source software, which can be applied to other cities and metropolitan areas, supporting the decarbonization of transport systems internationally.

>>>>

---
title: "Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from biclaR"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to public transport were also estimated and monetized.
  The results show that 10% of all trips could be made by cycling in combination with public transport. Shifting to cycling for the shorter first and last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the public transport leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

@@

We used 'jittering' to disaggregate the OD data, resulting in a wide spatial distribution of trip origins and destinations [@Lovelace2022Jittering].
The method works by sampling 'sub-points' (nodes on the transport network represented in OpenStreetMap in this case) and using these instead of a single point (typically the centroid) to represent trip origins and destinations for each zone.
This method then distributes the trips to desire lines connecting the subpoints based on a 'disaggregation threshold' which determines the maximum number of trips that can be represented by a single desire line.

@@

The jittering pre-processing stage generates a more realistic representation of the trips undertaken than the traditional centroid-based approach but does not precisely capture the exact spatial distribution of trips.
Even where such datasets exist, they cannot be shared for research due to data privacy regulations.

@@

<!-- Include a map with the legs of journeys, and which modes are replacing which, and what impacts are bein estimated at each one. See PTcar_legs.R -->

@@

For both *direct* and *safe* route profiles, 10% of the daily trips have the potential to be made by a combination of PT and cycling, even given the travel restrictions considered (up to 5 km on bike, up to 2 hours, no possible transfers between PT).
This unveils the potential of cycling as a complementary mode of PT, with the potential to uptake the number of PT trips within the LMA area by as much as 6.3% (in addition to the 825 thousand PT trips reported in the mobility survey).

@@

The quantification of benefits can support policy-makers in prioritizing interventions to reduce the reliance on private motorized modes of transportation.
The presentation of the results in an open access web application will help to inform and explain decisions.
Furthermore, the provision of datasets resulting from this project provides a foundation for further research and development of new tools and methods.
<!-- Punchy final sentence: -->
The methods are reproducible and based on open source software, which can be applied to other cities and metropolitan areas, supporting the decarbonization of transport systems internationally.

### 2023-11-30 paper/PaperCEUS/PaperCEUS.tex
last revision

  \begin{abstract}
  A high proportion of car trips can be replaced by a combination of
  public transit and cycling for the first-and-last mile. This paper
  estimates the potential for cycling combined with public transit as a
  substitute for car trips in the Lisbon metropolitan area and assesses
  its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of
  a metropolitan cycling network was developed (\emph{biclaR}). The
  social and environmental impacts were assessed using the \emph{HEAT
  for Cycling} and the \emph{HEAT as a Service} tools. The impacts of
  shifting car trips to public transport were also estimated and
  monetized. The results show that 10\% of trips could be made by
  cycling in combination with public transport. Shifting to cycling for
  the shorter first and last mile stages can reduce annual
  CO\textsubscript{2}eq emissions from 3,000 to 7,500 tons/day, while
  for the public transport leg, the transfer from car avoids of up to
  20,500 tons of CO\textsubscript{2}eq emissions per year. The estimated
  socio-environmental benefits are of €125 million to €325 million over
  10 years. This evidence can support policymakers to prioritize
  interventions that reduce the reliance on private motor vehicles.\\
  \end{abstract}
    \begin{keyword}
    Active transport \sep Intermodality \sep First and last
mile \sep Health economic assessment \sep Environmental impacts \sep 
    Open data and methods
  \end{keyword}

@@

We used `jittering' to disaggregate the OD data, resulting in a wide
spatial distribution trip origins and destinations (Lovelace et al.,
2022b). The method works by sampling `sub-points' (nodes on the
transport network represented in OpenStreetMap in this case) and using
these instead of a single point (typically the centroid) to represent
trip origins and destinations for each zone. This method then
distributes the trips to desire lines connecting the subpoints based on
a `disaggregation threshold' which determines the maximum number of
trips that can be represented by a single desire line.

@@

For both \emph{direct} and \emph{safe} route profiles, 10\% of the daily
trips have the potential to be made by a combination of PT and cycling
(up to 5 km on bike). This unveils the potential of cycling as a
complementary mode of PT, with the potential to uptake the number of PT
trips within the LMA area by as much as 6.3\% (in addition to the 825
thousand PT trips reported in the mobility survey).

@@

The quantification of benefits can support policy-makers in prioritizing
interventions to reduce the reliance on individual motorized
transportation. The presentation of the results in an open access web
application will help communicate decisions. Furthermore, the provision
of datasets resulting from this project provides a foundation for
further research and development of new tools and methods. The methods
are reproducible and based on open source software, which can be applied
to other cities and metropolitan areas, supporting the decarbonization
of transport systems internationally.

>>>>

  \begin{abstract}
  A high proportion of car trips can be replaced by a combination of
  public transit and cycling for the first-and-last mile. This paper
  estimates the potential for cycling combined with public transit as a
  substitute for car trips in the Lisbon metropolitan area and assesses
  its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of
  a metropolitan cycling network was developed (\emph{biclaR}). The
  social and environmental impacts were assessed using the \emph{HEAT
  for Cycling} and the \emph{HEAT as a Service} tools. The impacts of
  shifting car trips to public transport were also estimated and
  monetized. The results show that 10\% of all trips could be made by
  cycling in combination with public transport. Shifting to cycling for
  the shorter first and last mile stages can reduce annual
  CO\textsubscript{2}eq emissions from 3,000 to 7,500 tons/day, while
  for the public transport leg, the transfer from car avoids of up to
  20,500 tons of CO\textsubscript{2}eq emissions per year. The estimated
  socio-environmental benefits are of €125 million to €325 million over
  10 years. This evidence can support policymakers to prioritize
  interventions that reduce the reliance on private motor vehicles.\\
  \end{abstract}
    \begin{keyword}
    Active transport \sep Intermodality \sep First and last
mile \sep Health economic assessment \sep Environmental impacts \sep 
    Open data and methods
  \end{keyword}

@@

We used `jittering' to disaggregate the OD data, resulting in a wide
spatial distribution of trip origins and destinations (Lovelace et al.,
2022b). The method works by sampling `sub-points' (nodes on the
transport network represented in OpenStreetMap in this case) and using
these instead of a single point (typically the centroid) to represent
trip origins and destinations for each zone. This method then
distributes the trips to desire lines connecting the subpoints based on
a `disaggregation threshold' which determines the maximum number of
trips that can be represented by a single desire line.

@@

For both \emph{direct} and \emph{safe} route profiles, 10\% of the daily
trips have the potential to be made by a combination of PT and cycling,
even given the travel restrictions considered (up to 5 km on bike, up to
2 hours, no possible transfers between PT). This unveils the potential
of cycling as a complementary mode of PT, with the potential to uptake
the number of PT trips within the LMA area by as much as 6.3\% (in
addition to the 825 thousand PT trips reported in the mobility survey).

@@

The quantification of benefits can support policy-makers in prioritizing
interventions to reduce the reliance on private motorized modes of
transportation. The presentation of the results in an open access web
application will help to inform and explain decisions. Furthermore, the
provision of datasets resulting from this project provides a foundation
for further research and development of new tools and methods. The
methods are reproducible and based on open source software, which can be
applied to other cities and metropolitan areas, supporting the
decarbonization of transport systems internationally.
